Manuscript Figures

### 2024-04-01 manuscript_figures.Rmd
cleaned up directory

# Showing as a dendrogram
dend = as.dendrogram(hclust(d, method = "ward.D2"))

@@

@@

@@

```{r,fig.height = 11, fig.width = 8}
##Loading all the go stuff

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(shared_up_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_ECB.pdf",up_ECB,height=11,width=8)
```

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(shared_down_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_EcB.pdf",down_EcB,height =11, width = 18)
```
##### Faecalis

@@

```{r,fig.height = 11, fig.width = 8}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_faecalis.pdf",up_faecalis,height=11,width=8)
```

@@

```{r,fig.height = 11, fig.width=8}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(down_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_faecalis.pdf",down_faecalis,height=11,width=8)
```

@@

```{r,fig.height = 11, fig.width = 8}

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_faecium.pdf",up_faecium,height=11,width=8)
```

###### Down
```{r,fig.height = 11, fig.width = 8}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

@@

# Extracting the uniprotids up in infected and shared
down_in_faecium = infected_vs_healthy_plot %>% 
  filter(log2FC <= 0) %>% 
  filter(color == "sig_in_faecium") %>% 
  # Going from Protein ID to gene ID
  left_join(protein_to_gene,by = c("uniprotid" = "ProteinID")) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(down_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_faecium.pdf",down_faecium,height=11,width=8)

@@

```{r,fig.height = 11, fig.width= 8}
#Loading in data,
staph_data = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1)
anot_go = read_csv("01_r_processed_data/staph_go_annotations.csv") %>% 
  select(Entry = Gene.Names..primary.,Gene.Ontology.IDs) %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs))
    # Going from Protein ID to gene ID
 # left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>%

@@

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(sig_up,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

up_staph = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")
up_staph

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_staph.pdf",up_staph,height=11,width=8)
```

@@

```{r,fig.height = 11, fig.width= 8}
# Perfomring GO enrichment
enrich = clusterProfiler::enricher(sig_down,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_staph.pdf",down_staph,height=11,width=8)

@@

```{r,fig.height =10, fig.width = 12}

@@

# plotting the heatmap 
p1 = all %>% 
  ggplot(aes(reorder(name,size),type,fill = size))+
  geom_tile()+
  coord_flip()+
  scale_fill_gradient2(
  name = waiver(),
  high = "red",
  low = "green",
  mid = "white",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) + 
  xlab("GO term")+
  ylab("Bacteremia Type")+
  ggprism::theme_prism(base_size = 10)

@@

ggsave("02_figures/f2_infection_v_healthy_proteomics/f2b.pdf",height = 10, width = 12)

## ? - Venn Diagram

### Up in infected
```{r,fig.height = 3.8, fig.width=4}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

@@

# Generating venn diagram plot

library(RColorBrewer)
myCol <- brewer.pal(4, "Pastel2")

venn.diagram(
        x = list(enterococcus_sig, faecalis_sig, faecium_sig, staph_sig),
        category.names = c("EcB" , "E.faecalis " , "E.faecium","S. aureus"),
        filename = '02_figures/f2_infection_v_healthy_proteomics/up_in_infected_venn_diagram.png',
        output=TRUE,

        # Output features
        imagetype="png" ,
        height = 1100 , 
        width = 1100 , 
        resolution = 300,
        compression = "lzw",

)
```

@@

```{r,fig.height = 3.8, fig.width=4}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

@@

# Generating venn diagram plot

library(RColorBrewer)
myCol <- brewer.pal(4, "Pastel2")

venn.diagram(
        x = list(enterococcus_sig, faecalis_sig, faecium_sig, staph_sig),
        category.names = c("EcB" , "E.faecalis " , "E.faecium","S. aureus"),
        filename = '02_figures/f2_infection_v_healthy_proteomics/up_in_healthy_venn_diagram.png',
        output=TRUE,

        # Output features
        imagetype="png" ,
        height = 1100 , 
        width = 1100 , 
        resolution = 300,
        compression = "lzw",

)

```
## C- Evaluation of Proteomic Biomarkers

>>>>

# Showing as a dendrogram
dend = as.dendrogram(hclust(d,method = "ward.D2"))

@@

circlize_dendrogram(dend,
                    labels_track_height = NA,
                    dend_track_height = 0.5,
                    labels = TRUE)

```
# stats for dendrogram by k = 2 cluster
```{r,fig.height = 8, fig.width = 16}

circlize_dendrogram(dend %>% color_branches(k = 2),
                    labels_track_height = NA,
                    dend_track_height = 0.5,
                    labels = TRUE) 
stats 
```

@@

dend = as.dendrogram(hclust(d, "ward.D2"))

@@

```{r,fig.height = 10, fig.width = 10}
##Loading all the go stuff

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(shared_up_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

EcB_enrich_up = clusterProfiler::enricher(shared_up_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_ECB.pdf",up_ECB,height=10,width=10)
```

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(shared_down_in_infected,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_EcB.pdf",down_EcB,height =10, width = 10)
```
##### Faecalis

@@

```{r,fig.height = 10, fig.width = 10}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_faecalis.pdf",up_faecalis,height=10,width=10)
```

@@

```{r,fig.height = 10, fig.width=10}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(down_in_faecalis,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_faecalis.pdf",down_faecalis,height=10,width=10)
```

@@

```{r,fig.height = 10, fig.width = 10}

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(up_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_faecium.pdf",up_faecium,height=10,width=10)
```

###### Down
```{r,fig.height = 10, fig.width = 10}
#usng enricher
TERM2GENE = go %>% 
  dplyr::select(TERM = Gene.Ontology.IDs,GENE = Entry) %>% 
  as.data.frame()

@@

# Extracting the uniprotids up in infected and shared
down_in_faecium = healthy_v_faecium %>% 
  filter(log2FC >= 0) %>% 
  filter(adj.pvalue <= 0.05) %>% 
  pull(Gene)

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(down_in_faecium,
         universe = go$Entry,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_faecium.pdf",down_faecium,height=10,width=10)

@@

```{r,fig.height = 10, fig.width= 10}
#Loading in data,
staph_data = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1)
anot_go = read_csv("01_r_processed_data/staph_go_annotations.csv") %>% 
  select(Entry = Gene.Names..primary.,Gene.Ontology.IDs) %>% 
  dplyr::select(1,length(.)) %>% 
  tidyr::separate_rows(Gene.Ontology.IDs, sep = ';') %>% 
  dplyr::mutate(Gene.Ontology.IDs = gsub(" ","",Gene.Ontology.IDs))
    # Going from Protein ID to gene ID
 # left_join(protein_to_gene,by = c("Entry" = "ProteinID")) %>%

@@

test = TERM2GENE %>% 
  filter(TERM == "GO:0070062")

@@

# Perfomring GO enrichment
enrich = clusterProfiler::enricher(sig_up,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

  DT::datatable(enrich@result)

up_staph = clusterProfiler::cnetplot(enrich,
                                categorySize="pvalue",
                                cex_label_category = 1,
                                showCategory = 100,
                                cex_label_gene = 0.5)+
  theme(legend.position = "none")

up_staph

ggsave("03_supplemental/s2_networks_compared_to_healthy/up_staph.pdf",up_staph,height=10,width=10)
```

@@

```{r,fig.height = 10, fig.width= 10}
# Perfomring GO enrichment
enrich = clusterProfiler::enricher(sig_down,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME,
         maxGSSize = 1000)

DT::datatable(enrich@result)

# Perfomring GO enrichment
staph_down_enrich = clusterProfiler::enricher(sig_down,
         universe = background,
         TERM2GENE = TERM2GENE,
         TERM2NAME = TERM2NAME)

@@

ggsave("03_supplemental/s2_networks_compared_to_healthy/down_staph.pdf",down_staph,height=10,width=10)

@@

```{r,fig.height =10, fig.width = 7}

@@

# plotting the heatmap 
f2b = all %>% 
  ggplot(aes(reorder(name,size),type,fill = size))+
  geom_tile()+
  coord_flip()+
  scale_fill_gradient2(
  name = waiver(),
  high = "red",
  low = "green",
  mid = "white",
  midpoint = 0,
  space = "Lab",
  na.value = "grey50",
  guide = "colourbar",
  aesthetics = "fill"
) + 
  xlab("GO term")+
  ylab("Bacteremia Type")+
  ggprism::theme_prism(base_size = 10)+
  theme(legend.position = "top")

@@

ggsave("02_figures/f2_infection_v_healthy_proteomics/f2b.pdf",height = 10, width = 7)

### ? Platelet Alpha Graunle

```{r,fig.height = 5, fig.width = 10}

# Extracting the proteins with the term platelet alphaa granule lumen from staph enrichment results
staph_enrich_platelet = staph_down_enrich@result %>% 
  filter(Description == "platelet alpha granule lumen")

down_in_staph = strsplit(staph_enrich_platelet$geneID,"/")[[1]]

# Ex tracting the proteins with the term alph granule lumen from EcB enrichment results
EcB_enrich_platelet = EcB_enrich_up@result %>% 
  filter(Description == "platelet alpha granule lumen")

up_in_EcB = strsplit(EcB_enrich_platelet$geneID,"/")[[1]]

# These are the shared proteins at the intersection. 
shared = intersect(down_in_staph,up_in_EcB)
shared

# plotting differences to healthy - staph
staph = staph_data %>% 
  filter(Gene %in% down_in_staph) %>% 
  select(Gene,NN1:length(.)) %>% 
  pivot_longer(2:length(.),names_to = "sample_id",values_to = "Abundance") %>%
  mutate(Condition = case_when(grepl("HS.*|HM.*",sample_id) ~ "Infected",
                              TRUE ~ "Healthy"),
         Abundance = as.numeric(Abundance))

make_violin_plot_proteomics(staph,shared,ncol = length(shared),pname = "p")

# plotting differnces to healthy EcB 
protein_sum = read_csv("01_r_processed_data/quant_msstats_protein_level_data.csv")

f = protein_sum %>% 
  filter(Gene %in% shared) %>% 
  mutate(Condition = case_when(Condition == "Healthy" ~ "Healthy",
                               TRUE ~ "Infected"))

make_violin_plot_proteomics(f,shared,ncol = length(shared),pname = "p.signif")

## C - Venn Diagram

### Up in infected
```{r,fig.height = 5, fig.width=6}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

@@

library(ggVennDiagram)

up = ggVennDiagram(list(EcB = enterococcus_sig, faecalis = faecalis_sig, faecium = faecium_sig, S.aureus = staph_sig))+
  theme(legend.position = "none")+
  scale_fill_distiller(palette = "Reds", direction = 1)

up

ggsave("02_figures/f2_infection_v_healthy_proteomics/f2d_a.pdf", up, height = 5, width = 6)

ggVennDiagram(list(EcB = enterococcus_sig, S.aureus = staph_sig))

@@

```{r,fig.height = 5, fig.width=6}
infected_vs_healthy = readr::read_csv("01_r_processed_data/Infected_vs_Healthy.csv")
faecalis_v_healthy = readr::read_csv("01_r_processed_data/faecalis_v_healthy.csv")
healthy_v_faecium = readr::read_csv("01_r_processed_data/healthy_v_faecium.csv")

@@

library(ggVennDiagram)

down = ggVennDiagram(list(EcB = enterococcus_sig, faecalis = faecalis_sig, faecium = faecium_sig, S.aureus = staph_sig))+
  theme(legend.position = "none")+
  scale_fill_distiller(palette = "Greens", direction = 1)

down

ggsave("02_figures/f2_infection_v_healthy_proteomics/f2d_b.pdf",down,height = 5, width = 6)

ggVennDiagram(list(EcB = enterococcus_sig, S.aureus = staph_sig))

What proteins have differing responses in staph vs EcB?

```{r}
# found to be sig in both faecalis and faecium
enterococcus_sig_up = infected_vs_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC > 0)%>% 
  pull(uniprotid)

# Loading staph_data
staph_sig_down = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1) %>% 
  filter(Infection_pvalue <= 0.05,
         Infection_FC < 1) %>% 
  pull(Protein)

library(ggVennDiagram)
ggVennDiagram(list(enterococcus_sig_up, staph_sig_down))

GLabR::annotate_uniprot(intersect(enterococcus_sig_up,staph_sig_down),"accession,id,gene_primary,protein_name")
```

```{r}
# found to be sig in both faecalis and faecium
enterococcus_sig_down = infected_vs_healthy %>% 
  filter(adj.pvalue <= 0.05) %>% 
  filter(log2FC < 0)%>% 
  pull(uniprotid)

# Loading staph_data
staph_sig_up = readxl::read_excel("00_r_input_data/staph_proteomics.xlsx",skip = 1) %>% 
  filter(Infection_pvalue <= 0.05,
         Infection_FC > 1) %>% 
  pull(Protein)

library(ggVennDiagram)

ggVennDiagram(list(enterococcus_sig_down, staph_sig_up))

GLabR::annotate_uniprot(intersect(enterococcus_sig_down,staph_sig_up),"accession,id,gene_primary,protein_name")
```

## C- Evaluation of Proteomic Biomarkers
